\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-25 k-planar.tex
refs clean-up

\documentclass{patmorin}
\listfiles
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{amsfonts}
\usepackage{amsthm,mathtools}
\usepackage{paralist}
\usepackage{stmaryrd}
\usepackage{thm-restate}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\usepackage{todonotes}

\newcommand{\arXiv}[1]{arXiv:\,\href{https://arxiv.org/abs/#1}{#1}}
%\newcommand{\doi}[1]{\href{https://doi.org/#1}{\tt https://doi.org/#1}}

@@

\begin{proof}[Proof of \cref{d_framed_product_stucture}]
	Let $G$ be a $d$-framed multigraph with frame $G_0$ having outer face $F_0$.	Let $T$ be a BFS forest of $G_0$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.  Let $P_1:= V(F_0)$, let $Y_1$ be any subset of $\min\{|P_1|,3\}$ vertices in $P_1$ and let $X_1:=P_1\setminus Y_1$.  Now apply \cref{induction} on $G$, $G_0$, $T$, $F=F_0$, $k=1$, and $P_1=Y_1\cup X_1$.
    This gives an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ in which $H$ has a $3$-simple tree decomposition.  Since any graph having a $3$-simple tree decomposition is planar \cite{knauer.ueckerdt:simple,kratochvil.vaner:note,elmallah.colbourn:on}, this implies that $H$ is planar.

	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{\lfloor d/2\rfloor i}\cup \cdots \cup L_{\lfloor d/2\rfloor(i+1)-1}$ for each integer $i\ge 0$. That this is a layering of $G$ follows from the fact that $\dist_{G_0}(v,w)\le \lfloor d/2\rfloor$ for each edge $vw\in E(G)$.  Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\lfloor d/2\rfloor$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le \lfloor d/2\rfloor + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$. The result now follows from \cref{PartitionProduct}.
\end{proof}

@@

Since all the product structure theorems presented thus far upper bound the treewidth of $H$, \cref{qn} immediately implies bounds on the queue number of all graphs in these classes.  An improvement can be made for $1$-planar and $d$-map graphs since then $H$ is planar with treewidth at most $3$. \citet{ABGKP20} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$. Thus the graph $H$ in \cref{1_planar_product,dMapProduct} has queue-number at most $5$.
% , which with \cref{qn} implies:
% 
% \begin{cor}
% \label{1PlanarQueue}
% Every 1-planar graph has queue-number at most 
% $3 \times 7 \times 5 + \floor{\tfrac{3}{2} \times 7} = 115$.
% \end{cor}
% 
% \begin{cor}
% \label{dMapQueue}
% Every $d$-map graph has queue-number at most 
% $\floor{ \frac{33}{2} (d+3\floor{d/2}-3) }$.
% \note{DW}{I have updated this bound using the better qn bound for planar 3-trees. Note that with $d=3$ we get the original bound of 49.}
% \end{cor}
% 
% 
% 
% \cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$.
% \note{DW}{I have changed $g 2^{O(k^4)}$ to $g 2^{O(k^3)}$, okay?}
% \note{OM}{Ok}
% \note{DW}{I have changed $g>2^{k^3}$ to $g>2^{k^2}$, okay?}.
% \note{PM}{Ok}
The following corollary summarizes all of these results.
\begin{cor}\label{q_cor}
  The following bounds hold for the queue number of any graph from each of the following classes:
  \begin{compactitem}
    \item $k$-planar: $2^{O(k^3)}$
    \item $(g,k)$-planar: $\max\{1,g\}\cdot 2^{O(k^3)}$
    \item $(g,\delta)$-string: $\max\{1,g\}\cdot 2^{O(\delta^3)}$
    \item $1$-planar: $115$ 
    \item $d$-map: $\floor{ \frac{33}{2} (d+3\floor{d/2}-3) }$ 
    \item $(g,d)$-map: $\max\{1,g\}\cdot 2^{O(d)}$
    \item $k$-nearest-neighbour: $2^{O(k^6)}$
  \end{compactitem}
\end{cor}
%\note{PM}{For uniformity, I got rid of the two corollaries on $d$-map and $1$-planar graphs and kept them in \cref{q_cor}.  The text is still commented out in the source in case someone feels strongly about keeping them.}

@@

Combining \cref{non-repetitive} with our results on product structure, we obtain the following corollary:
\begin{cor}\label{nonrepetitive_cor}
  The following bounds hold for the nonrepetitive chromatic number of any graph from each of the following classes:
  \begin{compactitem}
    \item $k$-planar: $(18k^2+48k+30)\cdot 4^{\binom{k+4}{3}}$
    \item $(g,k)$-planar: $\max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}$
    % \item $(g,\delta)$-string: $g4^{O(\delta^3)}$
    \item $1$-planar: $1792$
    \item $d$-map: $256(d+3\floor{d/2}-3)$
    % \item $(g,d)$-map: $g4^{O(d)}$
    % \item $k$-nearest-neighbour: $4^{O(k^6)}$
    \item $(g,d)$-map graph: $(7d^2-21d)\cdot 4^{10}$
  \end{compactitem}
\end{cor}

>>>>

\documentclass{patmorin}
\listfiles
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{amsfonts}
\usepackage{amsthm,mathtools}
\usepackage{paralist}
\usepackage{stmaryrd}
\usepackage{thm-restate}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{todonotes}

@@

\begin{proof}[Proof of \cref{d_framed_product_stucture}]
	Let $G$ be a $d$-framed multigraph with frame $G_0$ having outer face $F_0$
	\note{DW}{Replace ``outer face'' by ``outer-cycle'' since a face is a defined to be a portion of the plane. }.	Let $T$ be a BFS forest of $G_0$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.  Let $P_1:= V(F_0)$, let $Y_1$ be any subset of $\min\{|P_1|,3\}$ vertices in $P_1$ and let $X_1:=P_1\setminus Y_1$.  Now apply \cref{induction} on $G$, $G_0$, $T$, $F=F_0$, $k=1$, and $P_1=Y_1\cup X_1$.
    This gives an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ in which $H$ has a $3$-simple tree decomposition.  Since any graph having a $3$-simple tree decomposition is planar \cite{knauer.ueckerdt:simple,kratochvil.vaner:note,elmallah.colbourn:on}, $H$ is planar.

	Define $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{\lfloor d/2\rfloor i}\cup \cdots \cup L_{\lfloor d/2\rfloor(i+1)-1}$ for each integer $i\ge 0$. This is a layering of $G$ since $\dist_{G_0}(v,w)\le \floor{ \frac{d}{2} }$ for each edge $vw\in E(G)$.  Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\floor{\frac{d}{2}}$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le \floor{ \frac{d}{2}} + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$. The result now follows from \cref{PartitionProduct}.
\end{proof}

@@

Since all the product structure theorems presented thus far upper bound the treewidth of $H$, \cref{qn} immediately implies bounds on the queue number of all graphs in these classes.  An improvement can be made for $1$-planar and $d$-map graphs since then $H$ is planar with treewidth at most $3$. \citet{ABGKP20} proved that every planar graph with treewidth at most $3$ has queue-number at most $5$. Thus the graph $H$ in \cref{1_planar_product,dMapProduct} has queue-number at most $5$.
% , which with \cref{qn} implies:
% 
% \begin{cor}
% \label{1PlanarQueue}
% Every 1-planar graph has queue-number at most 
% $3 \times 7 \times 5 + \floor{\tfrac{3}{2} \times 7} = 115$.
% \end{cor}
% 
% \begin{cor}
% \label{dMapQueue}
% Every $d$-map graph has queue-number at most 
% $\floor{ \frac{33}{2} (d+3\floor{d/2}-3) }$.
% \note{DW}{I have updated this bound using the better qn bound for planar 3-trees. Note that with $d=3$ we get the original bound of 49.}
% \end{cor}
% 
% 
% 
% \cref{qn,gkPlanarProduct} imply that $(g,k)$-planar graphs have queue-number at most $g 2^{O(k^3)}$.
% \note{DW}{I have changed $g 2^{O(k^4)}$ to $g 2^{O(k^3)}$, okay?}
% \note{OM}{Ok}
% \note{DW}{I have changed $g>2^{k^3}$ to $g>2^{k^2}$, okay?}.
% \note{PM}{Ok}
The following corollary summarizes all of these results.
\begin{cor}\label{q_cor}
  The following bounds hold for the queue number of any graph from each of the following classes:
  \begin{compactitem}
    \item $k$-planar: $2^{O(k^3)}$
    \item $(g,k)$-planar: $\max\{1,g\}\cdot 2^{O(k^3)}$
    \item $(g,\delta)$-string: $\max\{1,g\}\cdot 2^{O(\delta^3)}$
    \item $1$-planar: $115$ 
    \item $d$-map: $\floor{ \frac{33}{2} (d+3\floor{\frac{d}{2}}-3) }$ 
    \item $(g,d)$-map: $\max\{1,g\}\cdot 2^{O(d)}$
    \item $k$-nearest-neighbour: $2^{O(k^6)}$
  \end{compactitem}
\end{cor}
%\note{PM}{For uniformity, I got rid of the two corollaries on $d$-map and $1$-planar graphs and kept them in \cref{q_cor}.  The text is still commented out in the source in case someone feels strongly about keeping them.}

@@

Combining \cref{non-repetitive} with our results on product structure, we obtain the following corollary:
\begin{cor}\label{nonrepetitive_cor}
  The following bounds hold for the nonrepetitive chromatic number of any graph from each of the following classes:
  \begin{compactitem}
    \item $k$-planar: $(18k^2+48k+30)\cdot 4^{\binom{k+4}{3}}$
    \item $(g,k)$-planar: $\max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+4}{3}}$
    % \item $(g,\delta)$-string: $g4^{O(\delta^3)}$
    \item $1$-planar: $1792$
    \item $d$-map: $256(d+3\floor{\frac{d}{2}}-3)$
    % \item $(g,d)$-map: $g4^{O(d)}$
    % \item $k$-nearest-neighbour: $4^{O(k^6)}$
    \item $(g,d)$-map graph: $(7d^2-21d)\cdot 4^{10}$
  \end{compactitem}
\end{cor}
